---
title: "main"
output: pdf_document
date: "2024-06-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing dataset and libraries
```{r df_import, include=FALSE}

df= read.table("/Users/lorenzolaterza/Desktop/Data_Analysis_Final/BigMartSales.csv",header=TRUE, sep=",") 
library(dplyr)
library(corrplot)
library(ggplot2)
library(gridExtra)
```


Firstly, we are going to fix all features in our dataset, we will start from the feature "Item_Fat_Content".

We can see that there are different names for the same value:

```{r unique values}
item_fat_content_uniques <- unique(df$Item_Fat_Content)#unique values of item_fat_content

print(item_fat_content_uniques)
```
As we can see, we have 2 different names for "Regular Fat" that are "Regular" and "reg"
and 3 different names for "Low Fat" that are "Low fat", "low fat", "LF"

To solve this problem, we can edit the column "Item_Fat_Content" in this way:

all records that start with R or r, will be renamed "Regular Fat", and other will be renamed "Low Fat".

Then we will print the uniques name of the column after the edit

```{r sistemare nomi item_fat_content}
df$Item_Fat_Content <- ifelse(grepl("^[Rr]", df$Item_Fat_Content), "Regular Fat", "Low Fat")
item_fat_content_uniques <- unique(df$Item_Fat_Content)

print(item_fat_content_uniques)

```

Now we should work on the column "Outlet_Size" since analyzing the dataset manually it's clear that there are many missing values.
First of all, we should compute the percentage of missing values.

```{r Outlet size analisy}
#conteggio records senza Outlet_Size:
null<-sum(nchar(df$Outlet_Size) ==0)

#numero di records
total<- nrow(df)

print(paste("Number of rows with missing valuess on Outlet_Size:",null))

#percentuale di righe senza outlet_size
percentuale_stringhe_vuote <- (null / total) * 100

#stampa percentuale di righe senza outlet_size

print(paste("Percentuale di records con stringhe vuote in 'Outlet_Size':", percentuale_stringhe_vuote, "%"))
```

We want to find out if there are relations between the column "Outlet_Type" and the column "Outlet_Size" creating a table to see the number of combination between the outlet size and the outlet type.

```{r outlet_size and outlet_type}
# Creazione di una tabella di riepilogo
outlet_summary <- df %>%
  filter(Outlet_Size %in% c("Small", "Medium", "High")) %>%  # Filtra per includere solo le righe con i valori specificati
  group_by(Outlet_Type, Outlet_Size) %>%  # Raggruppa per tipo e dimensione del negozio
  summarise(Count = n(), .groups = 'drop')  # Calcola il conteggio e rimuove il raggruppamento automatico

# Visualizzazione della tabella di riepilogo
print(outlet_summary)
```
From this tab, we can see that all entries that are "Grocery Store" are small, and all entries that are "Supermarket Type2" or "Supermarket Type3" are Medium.
Knowing this, we can substitute blank values of "Outlet_Size" of grocery with small, of type2 and type 3 with medium.

```{r updating outlet_size}
# Aggiornamento della colonna 'Outlet_Size'
df <- df %>%
  mutate(Outlet_Size = case_when(
    Outlet_Type == "Grocery Store" ~ "Small",
    #Outlet_Type %in% c("Supermarket Type2", "Supermarket Type3") ~ "Medium",
    TRUE ~ Outlet_Size  # Mantiene il valore originale per tutte le altre condizioni
  ))

# Visualizza le modifiche per confermare
# Creazione di una tabella di riepilogo
outlet_summary <- df %>%
  filter(Outlet_Size %in% c("Small", "Medium", "High")) %>%  # Filtra per includere solo le righe con i valori specificati
  group_by(Outlet_Type, Outlet_Size) %>%  # Raggruppa per tipo e dimensione del negozio
  summarise(Count = n(), .groups = 'drop')  # Calcola il conteggio e rimuove il raggruppamento automatico

# Visualizzazione della tabella di riepilogo
print(outlet_summary)

```

```{r Outlet size analisy 2}
#conteggio records senza Outlet_Size:
null<-sum(nchar(df$Outlet_Size) ==0)

#numero di records
total<- nrow(df)

print(paste("Number of rows with missing valuess on Outlet_Size:",null))

#percentuale di righe senza outlet_size
percentuale_stringhe_vuote <- (null / total) * 100

#stampa percentuale di righe senza outlet_size

print(paste("Percentuale di records con stringhe vuote in 'Outlet_Size':", percentuale_stringhe_vuote, "%"))
```
Dopo aver riempito grocery store, supermarket type2 e supermarket type3, il numero di na è sceso dal 28% al 21%, passando da 2410 valori mancanti a 1855.

Now we will fill the remaining missing values with n/a
```{r adding na values}
# Aggiornamento della colonna 'Outlet_Size' per riempire le stringhe vuote
df <- df %>%
  mutate(Outlet_Size = if_else(nchar(Outlet_Size) == 0, "NA", Outlet_Size))
```


Now, we want to convert as factor the columns "Item_Fat_Content" giving 0 to "Low Fat" and 1 to "Regular".
And we also want to convert the column Outlet_Size:
  1->Small
  2->Medium
  3->Large

```{r factor converting}

# Conversione di 'Outlet_Size' in valori numerici
df <- df %>%
  mutate(Outlet_Size = case_when(
    Outlet_Size == "Small" ~ 1,
    Outlet_Size == "Medium" ~ 2,
    Outlet_Size == "High" ~ 3,
    TRUE ~ NA_real_  # Imposta NA per qualsiasi altro valore non specificato
  ))

df <- df%>%
  mutate(Item_Fat_Content = case_when(
    Item_Fat_Content == "Low Fat" ~ 1,
    Item_Fat_Content == "Regular Fat" ~ 2,
    TRUE ~ NA_real_
  ))
head(df)
```
```{r count 0 values}
# Find the number of zero 'Item_Visibility' values for each 'Outlet_Identifier'
zero_visibility_counts <- aggregate(Item_Visibility ~ Outlet_Identifier, data = df, function(x) sum(x == 0))

# Rename the column for better understanding
names(zero_visibility_counts)[2] <- "Zero_Item_Visibility_Count"

# Display the result
print(zero_visibility_counts)
```
```{r plots to see outliers}
# Load necessary libraries
library(ggplot2)
library(gridExtra)

# Histogram and Box Plot for Item_Weight
p1 <- ggplot(df, aes(x = Item_Weight)) + 
  geom_histogram(binwidth = 1, fill = 'blue', alpha = 0.7) + 
  ggtitle("Distribution of Item Weight") +
  xlab("Item Weight") +
  ylab("Frequency")

p2 <- ggplot(df, aes(x = "", y = Item_Weight)) + 
  geom_boxplot(fill = 'blue', alpha = 0.7) + 
  ggtitle("Box Plot of Item Weight") +
  xlab("") +
  ylab("Item Weight")

# Histogram and Box Plot for Item_Visibility
p3 <- ggplot(df, aes(x = Item_Visibility)) + 
  geom_histogram(binwidth = 0.01, fill = 'green', alpha = 0.7) + 
  ggtitle("Distribution of Item Visibility") +
  xlab("Item Visibility") +
  ylab("Frequency")

p4 <- ggplot(df, aes(x = "", y = Item_Visibility)) + 
  geom_boxplot(fill = 'green', alpha = 0.7) + 
  ggtitle("Box Plot of Item Visibility") +
  xlab("") +
  ylab("Item Visibility")

# Histogram and Box Plot for Item_MRP
p5 <- ggplot(df, aes(x = Item_MRP)) + 
  geom_histogram(binwidth = 5, fill = 'red', alpha = 0.7) + 
  ggtitle("Distribution of Item MRP") +
  xlab("Item MRP") +
  ylab("Frequency")

p6 <- ggplot(df, aes(x = "", y = Item_MRP)) + 
  geom_boxplot(fill = 'red', alpha = 0.7) + 
  ggtitle("Box Plot of Item MRP") +
  xlab("") +
  ylab("Item MRP")

# Histogram and Box Plot for Item_Outlet_Sales
p7 <- ggplot(df, aes(x = Item_Outlet_Sales)) + 
  geom_histogram(binwidth = 100, fill = 'purple', alpha = 0.7) + 
  ggtitle("Distribution of Item Outlet Sales") +
  xlab("Item Outlet Sales") +
  ylab("Frequency")

p8 <- ggplot(df, aes(x = "", y = Item_Outlet_Sales)) + 
  geom_boxplot(fill = 'purple', alpha = 0.7) + 
  ggtitle("Box Plot of Item Outlet Sales") +
  xlab("") +
  ylab("Item Outlet Sales")

# Print the plots individually
print(p1)
print(p2)
print(p3)
print(p4)
print(p5)
print(p6)
print(p7)
print(p8)

```

```{r count 0 values for each store}

visibility_sum_per_store <- aggregate(Item_Visibility ~ Outlet_Identifier, data = df, sum)
names(visibility_sum_per_store)[2] <- "Total_Item_Visibility"

# Display the result
print(visibility_sum_per_store)
```

```{r Visibility}
# Caricare il dataset

# Filtrare i record con visibilità maggiore di zero per calcolare le medie
filtered_df <- df[df$Item_Visibility > 0, ]

# Calcolare la media di Item_Visibility per ogni combinazione di Item_Type e Outlet_Size
visibility_avg_per_type_size <- aggregate(Item_Visibility ~ Item_Type + Outlet_Size, data = filtered_df, mean)

# Creare una chiave unica per facilitare il merge
visibility_avg_per_type_size$key <- with(visibility_avg_per_type_size, paste(Item_Type, Outlet_Size, sep = "_"))
df$key <- with(df, paste(Item_Type, Outlet_Size, sep = "_"))

# Merge tra i record del dataframe originale e le medie calcolate usando un left join
df <- merge(df, visibility_avg_per_type_size, by = "key", all.x = TRUE, suffixes = c("", ".new"))

# Sostituire i valori zero di Item_Visibility con i valori medi calcolati
df$Item_Visibility[df$Item_Visibility == 0] <- df$Item_Visibility.new[df$Item_Visibility == 0]

# Rimuovere le colonne in più create dal merge
df <- df[, !names(df) %in% c("Item_Visibility.new", "key")]

# Visualizzare il dataframe aggiornato
head(df)
```
```{r dropping useless columns}
df <- df %>%
  select(-Item_Type.new, -Outlet_Size.new)
```


```{r visibility plot}
# Histogram and Box Plot for Item_Visibility
p3 <- ggplot(df, aes(x = Item_Visibility)) + 
  geom_histogram(binwidth = 0.01, fill = 'green', alpha = 0.7) + 
  ggtitle("Distribution of Item Visibility") +
  xlab("Item Visibility") +
  ylab("Frequency")

p4 <- ggplot(df, aes(x = "", y = Item_Visibility)) + 
  geom_boxplot(fill = 'green', alpha = 0.7) + 
  ggtitle("Box Plot of Item Visibility") +
  xlab("") +
  ylab("Item Visibility")
print(p3)
print(p4)
```


```{r Removing Duplicates}
# Check for entire row duplicates
duplicates_entire_row <- duplicated(df)

# Count the number of entirely duplicated rows
num_entire_row_duplicates <- sum(duplicates_entire_row)
print(paste("Number of entirely duplicated rows:", num_entire_row_duplicates))

# View the duplicated rows
entire_row_duplicate_entries <- df[duplicates_entire_row, ]
print("Entirely duplicated rows:")
print(entire_row_duplicate_entries)
```

Now we will start to find out some correlations between variables.

```{R corr matrix}

# Calculate Spearman's rank correlation matrix again if not already calculated
cor_matrix <- cor(df %>% select(where(is.numeric)), 
                  method = "spearman", 
                  use = "pairwise.complete.obs")

# Visualize the correlation matrix with coefficients inside the circles
corrplot(cor_matrix, method = "circle", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, tl.cex = 0.8, cl.cex = 0.8,
         col = colorRampPalette(c("#6BAED6", "#FFFFFF", "#FD8D3C"))(200),
         addCoef.col = "black",  # Sets color of the coefficients to black (choose based on your color scheme)
         number.cex = 0.6)  # Adjust coefficient text size appropriately

```
DOBBIAMO DROPPARE LE COLONNE REMINAINING VISIBILITY EZERO ITEM VISIBILITY COUNT



We can see that there are big positive relation between:

outlet_size-item visibility
item_outlet_sales and outlet_size
item_mrp and item_outlet_sales

```{r item MRP vs Outlet Sales}
library(ggplot2)
ggplot(df, aes(x = Item_MRP, y = Item_Outlet_Sales)) +
  geom_point(alpha = 0.5) +
  labs(title = "Relationship between Item MRP and Outlet Sales",
       x = "Item MRP", y = "Outlet Sales")
```

```{r item MRP vs Outlet sales with gaps of price }
# Load necessary libraries
library(ggplot2)

# Specific price points to add to the plot
specific_price_points <- c(67.99, 134.49, 199.99)

# Plot the relationship between Item_MRP and Item_Outlet_Sales
p <- ggplot(df, aes(x = Item_MRP, y = Item_Outlet_Sales)) +
  geom_point(alpha = 0.5) +
  labs(title = "Relationship between Item MRP and Outlet Sales",
       x = "Item MRP", y = "Outlet Sales")

# Add vertical lines for specific price points
for (price in specific_price_points) {
  p <- p + geom_vline(xintercept = price, linetype = "dotted", color = "blue", alpha = 0.7)
  p <- p + annotate("text", x = price, y = max(df$Item_Outlet_Sales) * 0.9, label = price, color = "blue", angle = 90, vjust = -0.5, size = 3)
}

# Print the plot
print(p)

```

```{R item outlet sales by outlet size}
ggplot(df, aes(x = Outlet_Size, y = Item_Outlet_Sales)) +
  geom_boxplot(aes(fill = Outlet_Size)) +
  labs(title = "Item Outlet Sales by Outlet Size",
       x = "Outlet Size", y = "Outlet Sales")
```

```{R Sales trends over the years}
ggplot(df, aes(x = Outlet_Establishment_Year, y = Item_Outlet_Sales)) +
  geom_point(aes(color = Outlet_Establishment_Year)) +
  geom_smooth(method = "lm") +
  labs(title = "Sales Trends Over the Years",
       x = "Establishment Year", y = "Outlet Sales")

```

```{R Relationship between Item MRP and Outlet Sales by Product Type}

library(dplyr)

df_summary <- df %>%
  group_by(Item_Type, MRP_Bracket = cut(Item_MRP, breaks = seq(0, max(Item_MRP), by = 20))) %>%
  summarize(Average_Sales = mean(Item_Outlet_Sales), .groups = 'drop')

ggplot(data = df_summary, aes(x = MRP_Bracket, y = Average_Sales, fill = Item_Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Average Outlet Sales by MRP Bracket and Product Type",
       x = "MRP Bracket",
       y = "Average Outlet Sales") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ciaociao
#namoooo
#prova 3
```

zi qua c'è scritto che più vendi e più alzi il prezzo di base

```{R  Item_Outlet_Sales density plot}
# Create a density plot
ggplot(data = df, aes(x = Item_Outlet_Sales)) +
  geom_density(fill = "turquoise", alpha = 0.5) +  # 'alpha' controls transparency
  labs(title = "Density Plot of Item Outlet Sales",
       x = "Item Outlet Sales",
       y = "Density") +
  theme_light()

#ciao2

```

```{r Advanced visualization Visibility-Sales}
# Creating scatter plots segmented by Item_Type
p_item_type <- ggplot(df, aes(x = Item_Visibility, y = Item_Outlet_Sales)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  facet_wrap(~ Item_Type) +
  labs(title = "Item Visibility vs Outlet Sales by Item Type",
       x = "Item Visibility", y = "Outlet Sales") +
  theme_minimal()

# Creating scatter plots segmented by Outlet_Type
p_outlet_type <- ggplot(df, aes(x = Item_Visibility, y = Item_Outlet_Sales)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ Outlet_Type) +
  labs(title = "Item Visibility vs Outlet Sales by Outlet Type",
       x = "Item Visibility", y = "Outlet Sales") +
  theme_minimal()

# Print the plots
print(p_item_type)
print(p_outlet_type)
```

```{r Linear model to investigate on Item Visibility and Sales}
# Fit a linear model
model <- lm(Item_Outlet_Sales ~ Item_Visibility + Item_Type + Outlet_Type + Item_Fat_Content + Item_MRP, data = df)

# Summary of the model to understand influences
summary(model)
```

```{r Distribution of Variables across Item Sales}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Convert categorical variables to factor if not already
df$Item_Type <- as.factor(df$Item_Type)
df$Outlet_Type <- as.factor(df$Outlet_Type)
df$Item_Fat_Content <- as.factor(df$Item_Fat_Content)
df$Outlet_Size <- as.factor(df$Outlet_Size)
df$Outlet_Location_Type <- as.factor(df$Outlet_Location_Type)

# Continuous Variables
continuous_vars <- c("Item_Weight", "Item_Visibility", "Item_MRP")

# Categorical Variables
categorical_vars <- c("Item_Type", "Outlet_Type", "Item_Fat_Content", "Outlet_Size", "Outlet_Location_Type")

# Plotting Distribution of Sales for Continuous Variables
for(var in continuous_vars) {
  p <- ggplot(df, aes_string(x = var, y = "Item_Outlet_Sales")) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", color = "blue") +
    labs(title = paste("Scatter Plot of", var, "vs Outlet Sales"),
         x = var, y = "Outlet Sales") +
    theme_minimal()
  print(p)
}

# Plotting Distribution of Sales for Categorical Variables
for(var in categorical_vars) {
  p <- df %>%
    group_by_(.dots = var) %>%
    summarise(Average_Sales = mean(Item_Outlet_Sales)) %>%
    ggplot(aes_string(x = var, y = "Average_Sales", fill = var)) +
    geom_bar(stat = "identity", color = "black") +
    labs(title = paste("Average Sales by", var),
         x = var, y = "Average Sales") +
    theme_minimal()
  print(p)
}

```
